---
title: "Twitter Sentiment Analysis"
subtitle: "Tweets Hashtag: #SportsBetting"
author: "Andre Carney"
date: "June, 2022"
output: 
  word_document: 
    highlight: tango
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE, warning = FALSE)
```


<!-- ## Loading all the required packages in the R chunk below. -->

<!-- * You will need to install the libraries below if you don't have them install already. For instance, to install **tidyverse** from the console run **_install.packages("tidyverse")_** -->

```{r libraries}

library(rtweet)
#library(twitteR)
library(ROAuth)
require(RCurl)
library(stringr)
library(tm)
library(tidyverse)
library(wordcloud)
library(tidytext)
library(SnowballC)
library(fpc)
options(stringsAsFactors = FALSE, scipen = 999)

```


## Load Twitter API

* The first step is to register in the twitter application developer’s portal and get the authorization. You need to follow the link to register https://apps.twitter.com

* Create New App and use the API credentials in the project _.Renviron_ file. **Please do not display your API key in your publicly accessible code**.

```{r credentials}

# api_key = Sys.getenv("api_key")
# api_secret = Sys.getenv("api_secret")
# access_token = Sys.getenv("access_token")
# access_token_secret = Sys.getenv("access_token_secret")


```



 After obtaining credentials, we setup authorizations to access twitter API.

```{r credentialsKey}

#setup_twitter_oauth(api_key, api_secret, access_token, access_token_secret)

```



## Search twitter feeds

Search the twitter feeds  using the keyword below: 

* #SportsBetting 




### Mining twitter feeds with keyword: #SportsBetting

```{r searchkeyword}

## mutliple search keywords with hashtag
  
findfd <-   "#SportsBetting"

number <-  100000

```

 <!-- In the above code, we use keyword string to retrieve `r number ` tweets. The code for searching twitter is -->

```{r searchtwitter}
#tweets <- searchTwitter(findfd, number, lang = 'en', retryOnRateLimit = 100)

tweets <- search_tweets(
  findfd, n = number, include_rts = FALSE,lang = "en", 
  retryonratelimit = TRUE
)


n.tweet <- nrow(tweets)

cat("length of tweets retrieved is ", n.tweet)

```


<!-- ### Cleaning the tweets for further analysis and dropping all the retweets -->


```{r}

# tweets.df <- twListToDF(tweets) %>% 
#   filter(!isRetweet == "TRUE")

```


#### Text Analysis and Visualization

 <!-- Some data cleaning for descriptive visualizations starts from here -->

```{r}

tweet_data <- data.frame(date_time = tweets$created_at,
                         username = tweets$screen_name,
                         tweet_text = tweets$text,
                         favoritedCount = tweets$favorite_count,
                         retweetCount = tweets$retweet_count,
                         isRetweet = tweets$is_retweet,
                         retweeted_Count = tweets$retweet_retweet_count,
                         location = tweets$location)

# cleanup

tweet_data2 <- tweet_data %>% 
  mutate(date_time = as.POSIXct(date_time, format = "%a %b %d %H:%M:%S +0000 %Y")) %>%
  mutate(tweet_text = gsub("http://*| https://* | https)", "", tweet_text),
         tweet_text = gsub("https.*","",tweet_text),
         tweet_text = gsub("t.co", "", tweet_text),
         tweet_text = gsub('[[:digit:]]+', '', tweet_text),
         tweet_text = gsub('[[:punct:]]+', '', tweet_text))



### remove duplicated tweets

tweet_data2 <- tweet_data2[!duplicated(tweet_data2$tweet_text), ]

data("stop_words")

# get a list of words

tweet_clean <- tweet_data2 %>% 
  dplyr::select(tweet_text) %>%
  unnest_tokens(word, tweet_text) %>%
  filter(!(nchar(word) == 1)) %>% 
  anti_join(stop_words) %>%
  filter(!word %in% c("rt", "t.co"))

## joining, by = "word"

# plot the top 15 words --

tweet_clean_eng <- tweet_clean %>% 
         mutate(word = iconv(word, from = "latin1", to = "ASCII")) %>%
         filter(!is.na(word))

tweet_clean_eng$word <- str_replace_all(tweet_clean_eng$word, "[^[:alnum:]]", " ")


 tweet_clean_eng %>%
  count(word, sort = TRUE) %>%
  mutate(word = str_to_title(word),
         word = reorder(word, n)) %>% 
  top_n(15, n) %>%
   ungroup() %>% 
   arrange(word, -n) %>% 
  ggplot(aes(x = word, y = n, fill = word)) +
  geom_col(show.legend = F) +
   scale_y_continuous(labels = scales::comma)+
  xlab(NULL) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Words",
       title = "Top 15 words found in the tweets",
       caption = "tweets visualization | #BLM") +
   theme_bw()

```







WordCloud Visualization for significant words in the tweets

* Wordcloud showing the top 50 words. You can modify this in the code chunk below by adding the _top_n()_ function entry. 

```{r}

tweet_clean_eng <- tweet_clean %>% 
         mutate(word = iconv(word, from = "latin1", to = "ASCII")) %>%
         filter(!is.na(word))

tweet_clean_eng$word <- str_replace_all(tweet_clean_eng$word, "[^[:alnum:]]", " ")

wordCloud <- tweet_clean_eng %>% group_by(word) %>%
  summarise(Frequency = n()) %>% 
  top_n(50) #%>%  # top 50 words
#filter(!word %in% c("thepamilerin", "drmikeadenuga", "giveawaynaija", "governor",
                    #"makinde"))

wordcloud(wordCloud$word, wordCloud$Frequency, scale = c(5,.5), min.freq = 1, max.words = Inf,
	random.order = FALSE, random.color = TRUE, rot.per = .5,
	colors = "black" , ordered.colors = TRUE, use.r.layout = FALSE,
	fixed.asp = TRUE)

```


### Sentiment Analysis

<!-- This contains a lot of URLs, hashtags and other twitter handles. We will remove all these using the gsub function. -->

```{r}


tweets.df2 <- tweet_data %>% mutate(text = gsub("http://*| https://* | https)", "", tweet_text),
         text = gsub("https.*","",tweet_text),
         text = gsub("#.*", "", tweet_text),
         text = gsub("@.*", "", tweet_text))



```


 <!-- remove retweets and duplicate tweets text -->

```{r}
### remove retweeted tweets

tweets.df2 <- tweets.df2 %>% filter(isRetweet == FALSE)


### remove duplicated tweets

tweets.df2 <- tweets.df2[!duplicated(tweets.df2$tweet_text), ]

## remove non english character

tweets.df2$text <- str_replace_all(tweets.df2$text, "[[:punct:]]", " ")

tweets.df2 <- tweets.df2 %>% 
         mutate(text = iconv(text, to = "ASCII//TRANSLIT")) %>%
         filter(!is.na(text))



```




<!-- > Now, we have only the relevant part of the tweets and we can run our sentiment analysis part on the data. -->

### Getting sentiment score for each tweet

> We will first try to get the emotion score for each of the tweets using the _bing_ lexicon of **get_sentiments** function.

```{r}

tweets.df2_sentiment <- as.tibble(tweets.df2) %>%
  unnest_tokens(word, text) %>%
  inner_join(get_sentiments("bing"))


```




#### Visualize the overall sentiment

```{r}

result <- table(tweets.df2_sentiment$sentiment)

result2 <-  as.data.frame(result) 


result2 %>% 
  mutate(Var1 = factor(Var1, levels = c("positive", "negative")),
         Var1 = str_to_title(Var1)) %>% 
  group_by(Var1) %>%
  summarise(total = sum(Freq, na.rm = TRUE)) %>% 
  mutate(prop = total/sum(total)) %>% 
  filter(total != 0) %>% 
  
  ggplot(mapping = aes(x = 2, y = prop, fill = Var1))+
  geom_bar(width = 1, color = "white", stat = "identity") +
  xlim(0.5, 2.5) +
  coord_polar(theta = "y", start = 0) +
  theme_void() +
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(label = paste0(round(prop*100, 1), "%")), size = 3, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("#800020", "steelblue")) +
  labs(title = "Overall Sentiment",
       x = "",
       y = "",
       fill = "Category") +
  theme(legend.position = "top")

```




#### Plot the sentiment by keyword

```{r fig.height=5}
tweets.df2_sentiment %>% 
  group_by(word, sentiment) %>% 
  summarise(Count = n()) %>% 
  ungroup() %>% 
  #filter(Count > 1) %>% 
  top_n(50) %>% 
  ggplot(mapping = aes(x = reorder(word, Count), y = Count, fill = sentiment)) +
  geom_col(show.legend = FALSE, width = 0.5)+
  facet_wrap(.~ sentiment, scales = "free")+
  #theme(axis.text.x = element_text(angle = 90,size = 10, vjust = 0.5, hjust = 0.9))
  coord_flip() +
  theme_bw()+
  labs(y = "Frequency",
       x = "Terms")
  
```



## Topic Modelling using LDA

<!-- Additional libraries -->

```{r additional-libaries}
library(textmineR)
library(topicmodels)
```


### Data Preprocessing

```{r}
tweets.tbl <- tweets %>% 
  select(text, status_id)
tweets.tbl$text <- sub("RT.*:", "", tweets.tbl$text)
tweets.tbl$text <- sub("@.* ", "", tweets.tbl$text)
tweets.tbl$text <- sub("https", "", tweets.tbl$text)
tweets.tbl$text <- sub("t.co", "", tweets.tbl$text)
text_cleaning_tokens <- tweets.tbl %>% 
  tidytext::unnest_tokens(word, text)
text_cleaning_tokens$word <- gsub('[[:digit:]]+', '', text_cleaning_tokens$word)
text_cleaning_tokens$word <- gsub('[[:punct:]]+', '', text_cleaning_tokens$word)
text_cleaning_tokens <- text_cleaning_tokens %>% filter(!(nchar(word) == 1))%>% 
  anti_join(stop_words)
tokens <- text_cleaning_tokens %>% filter(!(word==""))
tokens <- tokens %>% mutate(ind = row_number())
tokens <- tokens %>% group_by(status_id) %>% mutate(ind = row_number()) %>%
  tidyr::spread(key = ind, value = word)
tokens [is.na(tokens)] <- ""
tokens <- tidyr::unite(tokens, text,-status_id,sep =" " )
tokens$text <- trimws(tokens$text)
```

### Create Document Term Matrix (DTM)
 
```{r}
#create DTM

dtm <- CreateDtm(tokens$text, 
                 doc_names = tokens$ID, 
                 ngram_window = c(1, 2))

#explore the basic frequency

tf <- TermDocFreq(dtm = dtm)
original_tf <- tf %>% select(term, term_freq,doc_freq)
rownames(original_tf) <- 1:nrow(original_tf)
```


### Modelling

```{r}
ap_lda <- LDA(dtm, k = 2, control = list(seed = 1234))
ap_topics <- tidytext::tidy(ap_lda, matrix = "beta") %>% 
  mutate(topic = case_when(topic == "1" ~ "Topic1",
                           topic == "2" ~ "Topic2"))

```


### Visualization

```{r}
ap_top_terms <- ap_topics %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  ungroup() %>% 
  arrange(topic, -beta)

ap_top_terms %>% 
  mutate(term = reorder(term, beta)) %>% 
  ggplot(aes(term, beta, fill = topic)) +
  geom_col(show.legend = FALSE) +
  facet_wrap( ~ topic, scales = "free") +
  coord_flip()
```

